---
title: "Graphic and Table Examples"
author: "Clifton Bell"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(ggplot2)
library(data.table)
library(stringr)
library(dplyr)
library(tidyr)
library(lubridate)
library(httr)
library(flextable)
library(jsonlite)
library(readr)
library(knitr)
library(kableExtra)
library(tibble)
library(gridExtra)
library(patchwork)
library(officer)

data_list <- readRDS(here("data", "data_list.rds"))
algal_taxa <- readRDS(here("data", "algal_taxa.rds"))

source(here("scripts", "data_import.R"))
source(here("scripts", "boxplots.R"))
source(here("scripts", "time_series.R"))
source(here("scripts", "algal_taxa.R"))
source(here("scripts", "scatterplots.R"))
source(here("scripts", "exceed_tables.R"))
source(here("scripts", "summary_stat_tables.R"))
```

This file contains examples of graphics and tables to support the Catawba River
reservoir water quality evaluation. I'm sharing to give Tom and Bill a flavor
of what we have to work with, which is quite a bit actually. Formatting of these
plots can be adjusted as needed, and stations and parameters can easily be 
swapped out.

I have not written any interpretation yet although a few things do pop out.
Overall I conclude that Lake Wateree shows the most signals of potential
eutrophication-related use impacts such as high pH, high cyanobacteria, T&O in
drinking water, etc. Upon coming back from vacation I plan on starting to write
interpretative report sections around these plots.

## Chlorophyll-a Boxplots

These plots help characterize the existing chla conditions in the reservoirs.
They only use growing season data from the upper meter of the water column. Only
stations with 5 or more data are used. The stations are plotted from upstream
to downstream.

```{r, echo=FALSE, fig.cap="Figure X - Spatial pattern in chl-a in Fishing Creek Reservoir."}
boxplots1("Chla", "FISHING CREEK RESERVOIR")
```

```{r, echo=FALSE, fig.cap="Figure X - Chl-a in Great Falls Reservoir."}
boxplots1("Chla", "GREAT FALLS RESERVOIR")
```

```{r, echo=FALSE, fig.cap="Figure X - Spatial pattern in chl-a in Cedar Creek Reservoir."}
boxplots1("Chla", "CEDAR CREEK RESERVOIR")
```

```{r, echo=FALSE, fig.cap="Figure X - Spatial pattern in chl-a in Lake Wateree."}
boxplots1("Chla", "LAKE WATEREE")
```

## Chlorophyll Time Series Plots

These plots would be used to infer trends in chla. Major trends are not apparent
in the charts. If there is more interest in trends we could perform a seasonal
Kendall trend test on selected stations.

```{r, echo=FALSE, fig.cap="Figure X - Chl-a over time at Fishing Creek Reservoir stations."}
time_series("Chla", c("CW-016F", "CW-057"))  #Fishing Creek Res.
```

```{r, echo=FALSE, fig.cap="Figure X - Chl-a over time at Fishing Creek Reservoir stations."}
time_series("Chla", c("CW-174", "CW-033"))  #Cedar Creek Res.
```

```{r, echo=FALSE, fig.cap="Figure X - Chl-a over time at Fishing Creek Reservoir stations."}
time_series("Chla", c("CW-231", "CW-208", "CL-089"))  #Lake Wateree
```

## Seasonal Plots for Chlorophyll-a, DO, and pH

These plots display the seasonal pattern in chlorophyll-a, DO, and pH. They only use growing
season data from the upper meter of the water column. Only stations with 5 or
more data are used.

```{r, echo=FALSE, fig.cap="Figure X - Seasonal pattern in chl-a at Fishing Creek Reservoir stations."}
boxplots2("Chla", c("CW-016F", "LCR-04", "CW-057"))   # Fishing Creek Res. Stations
```

```{r, echo=FALSE, fig.cap="Figure X - Seasonal pattern in chl-a at Cedar Creek Reservoir stations."}
boxplots2("Chla", c("CW-174", "CW-033"))              # Cedar Creek Res. Stations
```

```{r, echo=FALSE, fig.cap="Figure X - Seasonal pattern in chl-a at Lake Wateree Reservoir stations."}
boxplots2("Chla", c("CW-231", "CW-208", "CW-207B", "CL-089"))   # Lake Wateree Stations
```

```{r, echo=FALSE, fig.cap="Figure X - Seasonal pattern in DO at Fishing Creek Reservoir stations."}
boxplots2("DO", c("CW-016F", "LCR-04", "CW-057"), max_scale = 20)   # Fishing Creek Res. Stations
```

```{r, echo=FALSE, fig.cap="Figure X - Seasonal pattern in DO at Cedar Creek Reservoir stations."}
boxplots2("DO", c("CW-174", "CW-033"), max_scale = 20)              # Cedar Creek Res. Stations
```

```{r, echo=FALSE, fig.cap="Figure X - Seasonal pattern in DO at Lake Wateree Reservoir stations."}
boxplots2("DO", c("CW-231", "CW-208", "CW-207B", "CL-089"), max_scale = 20)   # Lake Wateree Stations
```

```{r, echo=FALSE, fig.cap="Figure X - Seasonal pattern in pH at Fishing Creek Reservoir stations."}
boxplots2("pH", c("CW-016F", "LCR-04", "CW-057"), min_scale = 5, max_scale = 10)   # Fishing Creek Res. Stations
```

```{r, echo=FALSE, fig.cap="Figure X - Seasonal pattern in pH at Cedar Creek Reservoir stations."}
boxplots2("pH", c("CW-174", "CW-033"), min_scale = 5, max_scale = 10)              # Cedar Creek Res. Stations
```

```{r, echo=FALSE, fig.cap="Figure X - Seasonal pattern in pH at Lake Wateree stations."}
boxplots2("pH", c("CW-231", "CW-208", "CW-207B", "CL-089"), min_scale = 5, max_scale = 10)   # Lake Wateree Stations
```

## Summary Tables for Chlorophyll-a, DO, pH, and Secchi depth

The following tables present summary statistics for chla, DO, pH, and Secchi depth
by reservoir and station. The stats are calculated from data from the growing season,
<= 1 m depth, and stations with at least five data.

```{r, echo = FALSE, results = 'asis'}
summary_table("Chla", "FISHING CREEK RESERVOIR")
summary_table("Chla", "GREAT FALLS RESERVOIR")
summary_table("Chla", "CEDAR CREEK RESERVOIR")
summary_table("Chla", "LAKE WATEREE")

summary_table("DO", "FISHING CREEK RESERVOIR")
summary_table("DO", "GREAT FALLS RESERVOIR")
summary_table("DO", "CEDAR CREEK RESERVOIR")
summary_table("DO", "LAKE WATEREE")

summary_table("pH", "FISHING CREEK RESERVOIR")
summary_table("pH", "GREAT FALLS RESERVOIR")
summary_table("pH", "CEDAR CREEK RESERVOIR")
summary_table("pH", "LAKE WATEREE")

summary_table("Secchi", "FISHING CREEK RESERVOIR")
summary_table("Secchi", "GREAT FALLS RESERVOIR")
summary_table("Secchi", "CEDAR CREEK RESERVOIR")
summary_table("Secchi", "LAKE WATEREE")
```

## Scatterplots of DO and pH vs Chlorophyll-a

The following plots help explore the relation between DO and pH and chlorophyll-a.
The regression lines can be turned off. The plots only use data <= 1 m.

```{r, echo=FALSE, fig.cap="Figure X - DO vs chla at Fishing Creek Reservoir stations.", results = "asis", message = FALSE, warning = FALSE}
scatter1("Chla", "DO", c("CW-016F", "LCR-04", "CW-057"), regression = TRUE)   # Fishing Creek Res. Stations
```

```{r, echo=FALSE, fig.cap="Figure X - DO vs chla at Great Falls Reservoir stations.", results = "asis", message = FALSE, warning = FALSE}
scatter1("Chla", "DO", c("RL-05414", "RL-08062", "RL-06429"), regression = TRUE)   # Great Falls Res. Stations
```

```{r, echo=FALSE, fig.cap="Figure X - DO vs chla at Cedar Creek Reservoir stations.", results = "asis", message = FALSE, warning = FALSE}
scatter1("Chla", "DO", c("CW-174", "CW-033"),  regression = TRUE)              # Cedar Creek Res. Stations
```

```{r, echo=FALSE, fig.cap="Figure X - DO vs chla at Lake Wateree Reservoir stations.", results = "asis", message = FALSE, warning = FALSE}
scatter1("Chla", "DO", c("CW-231", "CW-208", "CW-207B", "CL-089"),  regression = TRUE)   # Lake Wateree Stations
```

```{r, echo=FALSE, fig.cap="Figure X - pH vs chla at Fishing Creek Reservoir stations.", results = "asis", message = FALSE, warning = FALSE}
scatter1("Chla", "pH", c("CW-016F", "LCR-04", "CW-057"), regression = TRUE)   # Fishing Creek Res. Stations
```

```{r, echo=FALSE, fig.cap="Figure X - pH vs chla at Great Falls Reservoir stations.", results = "asis", message = FALSE, warning = FALSE}
scatter1("Chla", "pH", c("RL-05414", "RL-08062", "RL-06429"), regression = TRUE)   # Great Falls Res. Stations
```

```{r, echo=FALSE, fig.cap="Figure X - pH vs chla at Cedar Creek Reservoir stations.", results = "asis", message = FALSE, warning = FALSE}
scatter1("Chla", "pH", c("CW-174", "CW-033"),  regression = TRUE)              # Cedar Creek Res. Stations
```
```{r, echo=FALSE, fig.cap="Figure X - pH vs chla at Lake Wateree Reservoir stations.", results = "asis", message = FALSE, warning = FALSE}
scatter1("Chla", "pH", c("CW-231", "CW-208", "CW-207B", "CL-089"),  regression = TRUE)   # Lake Wateree Stations
```

## Scatterplots of DO vs Depth

Our report might not say anything about DO at depth, but these plots provide some
information on that topic.

```{r, echo=FALSE, fig.cap="Figure X - DO vs depth at Fishing Creek Reservoir stations.", results = "asis", message = FALSE, warning = FALSE}
scatter2(station_list = c("CW-016F", "LCR-04", "CW-057"))
```

```{r, echo=FALSE, fig.cap="Figure X - DO vs depth at Great Falls Reservoir stations.", results = "asis", message = FALSE, warning = FALSE}
scatter2(station_list = c("RL-05414", "RL-08062", "RL-06429"))
```

```{r, echo=FALSE, fig.cap="Figure X - DO vs depth at Cedar Creek Reservoir stations.", results = "asis", message = FALSE, warning = FALSE}
scatter2(station_list = c("CW-174", "CW-033"))
```

```{r, echo=FALSE, fig.cap="Figure X - DO vs depth at Lake Wateree stations.", results = "asis", message = FALSE, warning = FALSE}
scatter2(station_list = c("CW-231", "CW-208", "CW-207B", "CL-089"))
```

## DO and pH Exceedance Tables

These tables are intended to help interpret if the stations/reservoirs have
excessive DO (<5) or pH (> 8.5) violations. They only use data <= 1 m depth and
stations with at least five data. 


```{r, echo = FALSE, results='asis'}
below_table(param = "DO", threshold = 5, reservoir = "FISHING CREEK RESERVOIR")
below_table(param = "DO", threshold = 5, reservoir = "GREAT FALLS RESERVOIR")
below_table(param = "DO", threshold = 5, reservoir = "CEDAR CREEK RESERVOIR")
below_table(param = "DO", threshold = 5, reservoir = "LAKE WATEREE")

above_table(param = "pH", threshold = 8.5, reservoir = "FISHING CREEK RESERVOIR")
above_table(param = "pH", threshold = 8.5, reservoir = "GREAT FALLS RESERVOIR")
above_table(param = "pH", threshold = 8.5, reservoir = "CEDAR CREEK RESERVOIR")
above_table(param = "pH", threshold = 8.5, reservoir = "LAKE WATEREE")
```

## Algal Taxonomy Charts

These charts display both algal count and biovolume by major algal group. Chlorophyll-a
(<= 1 m) is displayed for comparison. All data were collected in 2017.

```{r, echo=FALSE, results = "asis", message = FALSE, warning = FALSE}
count_plot_16F <- stacked_algae(station_id = "CW-016F", metric = "count", start_date = "2017-06-01", end_date = "2017-10-31") 
biovolume_plot_16F <- stacked_algae(station_id = "CW-016F", metric = "biovolume", start_date = "2017-06-01", end_date = "2017-10-31")
chla_plot_16F <- plot_chla("CW-016F", start_date = "2017-06-01", end_date = "2017-10-31")
combined_16F_count <- (chla_plot_16F / count_plot_16F) 
combined_16F_biovol <- (chla_plot_16F / biovolume_plot_16F) 
count_plot_57 <- stacked_algae(station_id = "CW-057", metric = "count", start_date = "2017-06-01", end_date = "2017-10-31") 

biovolume_plot_57 <- stacked_algae(station_id = "CW-057", metric = "biovolume", start_date = "2017-06-01", end_date = "2017-10-31")
chla_plot_57 <- plot_chla("CW-057", start_date = "2017-06-01", end_date = "2017-10-31")
combined_57_count <- (chla_plot_57 / count_plot_57) 
combined_57_biovol <- (chla_plot_57 / biovolume_plot_57) 

count_plot_207 <- stacked_algae(station_id = "CW-207A", metric = "count", start_date = "2017-06-01", end_date = "2017-10-31") 
biovolume_plot_207 <- stacked_algae(station_id = "CW-207A", metric = "biovolume", start_date = "2017-06-01", end_date = "2017-10-31")
chla_plot_207 <- plot_chla("CW-207A", start_date = "2017-06-01", end_date = "2017-10-31")
combined_207_count <- (chla_plot_207 / count_plot_207) 
combined_207_biovol <- (chla_plot_207 / biovolume_plot_207) 

count_plot_89 <- stacked_algae(station_id = "CL-089", metric = "count", start_date = "2017-06-01", end_date = "2017-10-31") 
biovolume_plot_89 <- stacked_algae(station_id = "CL-089", metric = "biovolume", start_date = "2017-06-01", end_date = "2017-10-31")
chla_plot_89 <- plot_chla("CL-089", start_date = "2017-06-01", end_date = "2017-10-31")
combined_89_count <- (chla_plot_89 / count_plot_89) 
combined_89_biovol <- (chla_plot_89 / biovolume_plot_89) 

# Your existing code for creating plots

# Print plots with captions
print_with_caption(combined_16F_count, "Figure X: CW-016F (Fishing Creek Res.) Chlorophyll-a and Algae Count")
print_with_caption(combined_16F_biovol, "Figure X: CW-016F  (Fishing Creek Res.) Chlorophyll-a and Algae Biovolume")
print_with_caption(combined_57_count, "Figure X: CW-057  (Fishing Creek Res.) Chlorophyll-a and Algae Count")
print_with_caption(combined_57_biovol, "Figure X: CW-057  (Fishing Creek Res.) Chlorophyll-a and Algae Biovolume")
print_with_caption(combined_207_count, "Figure X: CW-207A (Lake Wateree) Chlorophyll-a and Algae Count")
print_with_caption(combined_207_biovol, "Figure X: CW-207A (Lake Wateree) Chlorophyll-a and Algae Biovolume")
print_with_caption(combined_89_count, "Figure X: CL-089 (Lake Wateree) Chlorophyll-a and Algae Count")
print_with_caption(combined_89_biovol, "Figure X: CL-089 (Lake Wateree) Chlorophyll-a and Algae Biovolume")
```

## Cyanobacteria Counts and WHO Risk Thresholds

These charts compare the 2017 cyanobacteria counts to WHO risk thresholds


```{r, echo=FALSE, results = "asis", message = FALSE, warning = FALSE}
cyano_16F <- cyano(station_id = "CW-016F", start_date = "2017-06-01", end_date = "2017-10-31")
cyano_combined_16F <- (chla_plot_16F / cyano_16F)

cyano_57 <- cyano(station_id = "CW-057", start_date = "2017-06-01", end_date = "2017-10-31")
cyano_combined_57 <- (chla_plot_57 / cyano_57)

cyano_207 <- cyano(station_id = "CW-207A", start_date = "2017-06-01", end_date = "2017-10-31")
cyano_combined_207 <- (chla_plot_207 / cyano_207)

cyano_89 <- cyano(station_id = "CL-089", start_date = "2017-06-01", end_date = "2017-10-31")
cyano_combined_89 <- (chla_plot_89 / cyano_89)

print_with_caption(cyano_combined_16F, "Figure X: Cyanobacteria counts at station CW-16F (Fishing Creek Reservoir).")
print_with_caption(cyano_combined_57, "Figure X: Cyanobacteria counts at station CW-057 (Fishing Creek Reservoir).")
print_with_caption(cyano_combined_207, "Figure X: Cyanobacteria counts at station CW-207A (Lake Wateree).")
print_with_caption(cyano_combined_89, "Figure X: Cyanobacteria counts at station CL-089 (Lake Wateree).")
```

## Scatterplot of Cyanobacteria Counts vs. Chlorophyll-a

These plots are intended to explore the relation between cyanobacteria cell counts
and chlorophyll-a.

```{r FC_version, echo=FALSE, results = "asis", message = FALSE, warning = FALSE, fig.cap = "Figure X: Cyanobacteria count vs. chla, Fishing Creek Reservoir"}
chla_mean <- data_list$Chla %>%
  filter(Waterbody == "FISHING CREEK RESERVOIR", Depth_m <= 1) %>%
  group_by(StationID, ActivityStartDate) %>%
  summarize(mean_Chla = mean(Chla, na.rm = TRUE)) %>%
  ungroup()

algal_summarized <- algal_taxa %>%
  filter(Waterbody == "FISHING CREEK RESERVOIR", Algal_Group == "Cyanobacteria") %>%
  group_by(StationID, ActivityStartDate) %>%
  summarize(total_Cells = sum(Cells, na.rm = TRUE)) %>%
  ungroup()

merged_data <- inner_join(chla_mean, algal_summarized, 
                          by = c("StationID", "ActivityStartDate"))

ggplot(merged_data, aes(x = mean_Chla, y = total_Cells, color = StationID)) +
  geom_point() +
  geom_hline(yintercept = 20000, linetype = "dashed", color = "blue") +
  geom_hline(yintercept = 100000, linetype = "dashed", color = "blue") +
  annotate("text", x = max(merged_data$mean_Chla), y = 20000, 
           label = "WHO moderate risk threshold", hjust = 1, vjust = -0.5, color = "blue") +
  annotate("text", x = max(merged_data$mean_Chla), y = 100000, 
           label = "WHO high risk treshold", hjust = 1, vjust = -0.5, color = "blue") +
  labs(x = "Mean Chlorophyll-a", y = "Total Cyanobacteria Cells",
       title = "Total Cyanobacteria Cells vs Mean Chlorophyll-a in Fishing Creek Reservoir",
       color = "Station ID") +
  theme_minimal() +
  scale_y_continuous(labels = scales::comma)
```

```{r LW_version, echo=FALSE, results = "asis", message = FALSE, warning = FALSE, fig.cap = "Figure X: Cyanobacteria count vs. chla, Lake Wateree"}
chla_mean <- data_list$Chla %>%
  filter(Waterbody == "LAKE WATEREE", Depth_m <= 1) %>%
  group_by(StationID, ActivityStartDate) %>%
  summarize(mean_Chla = mean(Chla, na.rm = TRUE)) %>%
  ungroup()

algal_summarized <- algal_taxa %>%
  filter(Waterbody == "LAKE WATEREE", Algal_Group == "Cyanobacteria") %>%
  group_by(StationID, ActivityStartDate) %>%
  summarize(total_Cells = sum(Cells, na.rm = TRUE)) %>%
  ungroup()

merged_data <- inner_join(chla_mean, algal_summarized, 
                          by = c("StationID", "ActivityStartDate"))

ggplot(merged_data, aes(x = mean_Chla, y = total_Cells, color = StationID)) +
  geom_point() +
  geom_hline(yintercept = 20000, linetype = "dashed", color = "blue") +
  geom_hline(yintercept = 100000, linetype = "dashed", color = "blue") +
  annotate("text", x = max(merged_data$mean_Chla), y = 20000, 
           label = "WHO moderate risk threshold", hjust = 1, vjust = -0.5, color = "blue") +
  annotate("text", x = max(merged_data$mean_Chla), y = 100000, 
           label = "WHO high risk threshold", hjust = 1, vjust = -0.5, color = "blue") +
  labs(x = "Mean Chlorophyll-a", y = "Total Cyanobacteria Cells",
       title = "Total Cyanobacteria Cells vs Mean Chlorophyll-a in Lake Wateree",
       color = "Station ID") +
  theme_minimal() +
  scale_y_continuous(labels = scales::comma)
```

# Percent Cynaobacteria Biovolume vs Chlorophyll-a

These plots are intended to explore the relation between percent cyanobacteria
biovolume and chlorophyll-a.


```{r FC_version2, echo=FALSE, results = "asis", message = FALSE, warning = FALSE, fig.cap = "Figure X: Cyanobacteria biovolume vs. chla, Fishing Creek Reservoir"}

chla_mean <- data_list$Chla %>%
  filter(Waterbody == "FISHING CREEK RESERVOIR", Depth_m <= 1) %>%
  group_by(StationID, ActivityStartDate) %>%
  summarize(mean_Chla = mean(Chla, na.rm = TRUE)) %>%
  ungroup()

algal_summarized <- algal_taxa %>%
  filter(Waterbody == "FISHING CREEK RESERVOIR", Algal_Group == "Cyanobacteria") %>%
  group_by(StationID, ActivityStartDate) %>%
  summarize(Sum_Percent_Biovol = sum(Perc_Biovol, na.rm = TRUE)) %>%
  ungroup()

merged_data <- inner_join(chla_mean, algal_summarized, 
                          by = c("StationID", "ActivityStartDate"))

ggplot(merged_data, aes(x = mean_Chla, y = Sum_Percent_Biovol, color = StationID)) +
  geom_point() +
  labs(x = "Mean Chlorophyll-a", y = "Percent Cyanobacteria Biovolume",
       title = "Percent Cyanobacteria Biovolume vs Mean Chlorophyll-a in Fishing Creek Reservoir",
       color = "Station ID") +
  theme_minimal() +
  scale_y_continuous(labels = scales::comma)
```

```{r LW_version2, echo=FALSE, results = "asis", message = FALSE, warning = FALSE, fig.cap = "Figure X: Cyanobacteria biovolume vs. chla, Lake Wateree"}
chla_mean <- data_list$Chla %>%
  filter(Waterbody == "LAKE WATEREE", Depth_m <= 1) %>%
  group_by(StationID, ActivityStartDate) %>%
  summarize(mean_Chla = mean(Chla, na.rm = TRUE)) %>%
  ungroup()

algal_summarized <- algal_taxa %>%
  filter(Waterbody == "LAKE WATEREE", Algal_Group == "Cyanobacteria") %>%
  group_by(StationID, ActivityStartDate) %>%
  summarize(Sum_Percent_Biovol = sum(Perc_Biovol, na.rm = TRUE)) %>%
  ungroup()

merged_data <- inner_join(chla_mean, algal_summarized, 
                          by = c("StationID", "ActivityStartDate"))

ggplot(merged_data, aes(x = mean_Chla, y = Sum_Percent_Biovol, color = StationID)) +
  geom_point() +
  labs(x = "Mean Chlorophyll-a", y = "Percent Cyanobacteria Biovolume",
       title = "Percent Cyanobacteria Biovolume vs Mean Chlorophyll-a in Lake Wateree",
       color = "Station ID") +
  theme_minimal() +
  scale_y_continuous(labels = scales::comma)
```



